---
title: "Central Limit Theorem - An Illustration"
output: pdf_document
---

```{r preamble, include=FALSE}
library(ggplot2)
library(reshape2)
library(dplyr)
library(latex2exp)

rm(list = ls())


```

## Introduction
This document is a rudimentary illustration of the Central Limit Theorem. We start with an exponential distribution, and observe the distribution of sample means.


## Theory
Let $X$ follow an exponential distribution with rate $\lambda$, thus

\begin{eqnarray*}
\text{E}(X) &=& \frac{1}{\lambda} \\
\text{Var}(X) &=& \frac{1}{\lambda^2} \text{, and} \\
\sigma_{X} &=& \sqrt{\text{Var}(X)} = \frac{1}{\lambda}
\end{eqnarray*}

Let $\left\{X_i\right\}_{i=1}^{n}$ be a series of $n$ i.i.d. variables s.t. $X_i\sim\text{exp}(\lambda)$. We define the sample mean $\bar{X} = \frac{1}{n}\sum_{i=1}^{n}X_i$. We can see that

\begin{eqnarray*}
\text{E}(\bar{X}) &=& \frac{1}{n}\sum_{i=1}^{n}\text{E}(X_i) = \frac{1}{\lambda}\\
\text{Var}(\bar{X}) &=& \sum_{i=1}^{n}\frac{1}{n^2}\text{Var}(X_i)=\frac{1}{n\lambda^2}\text{, and} \\
\sigma_{\bar{X}} &=& \sqrt{\text{Var}(\bar{X})} = \frac{1}{\sqrt{n}\lambda}
\end{eqnarray*}

Thus, according to the Central Limit Theorem, when $n$ is sufficiently large, $\bar{X}$ should converge towards a normal distribution with mean $\frac{1}{\lambda}$ and standard deviation $\frac{1}{\sqrt{n}\lambda}$.



## Simulation
For the purpose of simulation, we set $\lambda = 0.20$, $n = 40$, we run the simulation $1,000$ times.

```{r}


lambda <- 0.2
n_sims <- 1000
n <- 40


theo_x <- seq(2.5, 7.5, length = 1000)

theo_df<- data.frame(
        x = theo_x,
        y = dnorm(theo_x, mean = 1/lambda, sd = 1/(lambda*sqrt(n)))
)

rm(theo_x)


set.seed(100)
sim_data <- matrix(rexp(n_sims * n, rate = lambda), nrow = n_sims)


```





```{r}

lambda <- 0.2
n_sims <- 1000
n <- 40

sim_data <- matrix(rexp(n_sims * n, rate = lambda), nrow = n_sims)

means_df <- data.frame(
        x = apply(sim_data, 1, mean)
)

x_bar <- apply(sim_data, 1, mean)
x_bar_mean <- mean(x_bar)
x_bar_sd <- sd(x_bar)
```

In theory, $\bar{X}$ should have a mean of $\frac{1}{\lambda} = 5$, and standard deviation of $\frac{1}{\sqrt{n}\lambda}$ = `r 1/(sqrt(n) * lambda)`. We observe a sample mean of `r x_bar_mean` and standard deviation of `r x_bar_sd`.


## Plotting

Visually, the sample means follow a distribution which is approximately normal. There are more sophisticated ways to illustrate the convergence (e.g., Q-Q plots, comparing moments, etc.) but the following chart would suffice for our purposes.


```{r}
g <- ggplot(means_df, aes(x = x, y = ..density.., fill = "Simulated")) +
        geom_histogram(bins = 50) +
        geom_line(data = theo_df,
                  aes(y = y, color = "Theoretical")) +
        scale_color_manual(name = NULL,
                           values = c("Theoretical" = "blue")) +
        scale_fill_manual(name = NULL,
                          values = c("Simulated" = "grey")) +
        theme(legend.position = "top") +
        ggtitle(TeX("Simulated v. Theoretical Distribution of $\\bar{X}$"),
                subtitle = TeX("where \\bar{X}=\\frac{1}{40}\\sum_{i=1}^{40}X_i, and X_1, X_2, ..., X_{40} are iid and follow an exponential distribution with $\\lambda = 0.2$"))
g
```

